\section{Applications}
\subsection{AdvertService Adaptor}
\label{advertservice}
Our first implementation of our Advert server is probably the most obvious use
of the Advert server; the \emph{JavaGAT AdvertService}. The AdvertService
allows \emph{Advertisable} instances to get published to and queried in an
advert directory. Such an advert directory is a meta data directory with an
hierarchical namespace attached \cite{javagat-javadoc}.
\subsubsection{Adaptor Implementation}
For the actual implementation of our \emph{AppEngineAdvertServiceAdaptor}, we
basically implemented all functionality provided by the JavaGAT API
\cite{javagat-javadoc}. This makes our implementation very similar to the
\emph{GenericAdvertServiceAdaptor}, with the major difference that we do not
use a local database for storage. Instead, we make use of the \emph{Ibis Advert
Client} library, which communicates with the App Engine (see subsection 
\ref{clientimpl}). 

In the adaptor's constructor we connect to the App Engine by creating a new
\texttt{Advert} object (which in turn connects to the App Engine using the
\texttt{Communications} class). The constructor only takes one argument, namely
an \texttt{org.gridlab.gat.URI} of the Advert Server to connect to. This
\texttt{URI} needs to be an \texttt{appspot.com} domain. Subsequently, the
constructor tries to fetch the user's credentials from the \texttt{GATContext}. 
If none is found, we will try to connect to the Advert server as a public
server.

Because of the fact that we do not implement our database locally, we cannot
implement two of the public functions given by the API, called
\texttt{importDataBase} and \texttt{exportDataBase}. Both functions are not
feasible to implement, since this would mean we theoratically should be able to
import/export 1 Gigabyte of data over HTTP, after which we would run out of
free quota too fast.

\subsection{IPL Registry Bootstrap}
\label{ipl}
Besides implementing a AdvertService adaptor for the Google App Engine, we
found another use for our advert service library, within Ibis. Our advert
server could very well function as a bootstrap server for an \texttt{IPL} (Ibis
Portability Layer) \cite{ipl-www} registry.

\subsubsection{IPL Server}
The Ibis Portability layer is a communication library is specifically
designed for usage in a grid environment. It has a number of properties which
help to achieve its goal of providing programmers with an easy to use, reliable
grid communication infrastructure \cite{ipl-www}.

A central concept in Ibis is the \emph{pool}. A pool consists of one or more Ibis
instances, usually running on different machines. Each pool is generally made up
of Ibises running a single distributed application. Ibises in a pool can
communicate with each other, and, using the registry mechanism present in Ibis,
can search for other Ibises in the same pool, get notified of Ibises joining the
pool, etc. To coordinate Ibis pools a socalled \emph{Ibis server} is used
\cite{ipl-usersguide}.

Services can be dynamically added to the server. By default, the Ibis
communication library comes with a \emph{registry service}. This registry
service manages pools, possibly multiple pools at the same time. Before starting
an Ibis application, an Ibis server needs to be running on a machine that is
accessible from all nodes participating in the Ibis run. Starting such
application requires the user to specify the server's address to the
application, in order to join a pool. Specifying such address can be well done
by means of a \emph{Registry Bootstrap Service}, in which addresses are
specified, described by an identifier and (optional) meta data.

\subsubsection{Registry Bootstrap Service (Server)}
Mapping our requirements of an IPL Registry Bootstrap service on our existing
Advert service, we have to make some design decisions. First of all, the user
should be able to decide whether a service is to be advertised or not. This can
be specified once the server is started, by using an extra argument when calling
\texttt{ipl-server}. For example, we could add \texttt{--advert ADVERT\_URI} as
parameter, after which we store the server's address at a certain advert service
, under a specific identifier (indicated by \texttt{ADVERT\_URI}). We structured
the \texttt{ADVERT\_URI} as follows:

\begin{center}
\begin{code}
google://jondoe\%40gmail.com:password@jondoe.appspot.com/some_identifier
\end{code}
\end{center}

In this schema \texttt{google://} refers to the Goole App Engine. Next we can
optionally add the user's credentials (there are two ways to do this, as we
will see below). To authenticate to the Advert server, a Google Account (in
the form of an email address is required). Since the \texttt{@} symbol is
reserved as identifier for the URI (RFC 2396), we have to use an URI-encoded
\texttt{@} symbol, which is \texttt{\%40}. After the actual \texttt{@}
symbol, the hostname is presented. The server does not necessarily need to run
on Google's \texttt{appspot.com} domain, but for now, we assume it will.
Everyting from the trailing \texttt{/} and after is considered the identifier at
which the IPL Registry is stored.

Alternatively, it is also possible for the user to provide its credentials
using a seperate commandline command. The user can specify both \texttt{--user
USERNAME} and \texttt{--pass PASSWORD}, followed by its username and password
respectively. If the credentials are specified twice (i.e. using both the
\texttt{ADVERT\_URI} and the \texttt{--user} and \texttt{--pass} to specify
username and password), preference is given to the second pair of
username/password, and a warning is issued.

Secondly, we have to find a useful way to make use of meta data. Storing
key-value pairs can be really useful, if structured properly. Some examples
would be an \texttt{author} field (the user that started the IPL server), a
field with some sort of \texttt{time-of-creation} (supplying the time the
server was created), etc. We still need a means to properly structure
and pass meta data as soon as the server is created. The easiest way is to pass
an additional parameter like \texttt{--metadata}, after which we specify our
meta data as follows:

\begin{center}
\begin{code}
key1=value1,key2=value2,key3=value3
\end{code}
\end{center}

This way, our meta data is almost specified as it would be in JSON, which makes
it relatively easy to convert to a \texttt{MetaData} Object. No spaces or
special characters (i.e. \texttt{=} or \texttt{,}) are allowed. Optionally we
could implement meta data being present in a separate file, which would require a
file parser of some kind. This is a future feature to implement.

Once a server is started using the \texttt{--advert} suffix, the server is added
to the Advert server (using the \texttt{add()} functionality already present).
Once a server is stopped, it will be removed (using the \texttt{del()}
function).

\subsubsection{Ibis Application (Client)}
Subsequently, when we start an Ibis application to join a certain pool, and we
are not aware of the server address, we could specify
\texttt{-Dibis.advert.address} (where the address is structured like the
\texttt{ADVERT\_URI} as stated above), instead of specifying
\texttt{-Dibis.server.address}. Other properties that can be specified are:

\begin{itemize}
  \item \texttt{-Dibis.advert.username}; Username, if preferred above
  	\texttt{ibis.advert.address}
  \item \texttt{-Dibis.advert.password}; Password, if preferred above
  	\texttt{ibis.advert.address} 
  \item \texttt{-Dibis.advert.metadata}; Meta data, if no server identifier is
  	specified in \newline \texttt{ibis.advert.address}
\end{itemize}

Again, if the user's credentials are specified twice, preference is given to
\texttt{ibis.advert.username} / \texttt{password}, and a warning is issued.
Furthermore, if no Advert id is given in the \texttt{ADVERT\_URI}, meta data
could be used to retrieve an appropriate registry server. If more servers
adhere to a particular set of meta data, the first one returned by the Advert
service is used.

Once the right registry server is retrieved, the \texttt{ibis.server.address}
property is set, so an application can connect to a suitable registry server as
it would normally do, if \texttt{-Dibis.server.address} was set in the first
place.

